DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

\subsection{}
\begin{mdframed}[style=comment]
In Appendix C. is there a connection between $t$ and $\sigma$ that you can make, or do you just fit them to be similar? If you just fit them to be similar I'm not sure where the relative differences come from. If they are related, however, then the relative differences make sense, but I can't see what the connection is.
\end{mdframed}

In the current version of the manuscript and the code, we simply fitted a few $t$ to match the different $\sigma$.
In theory, there should be a connection between the two values providing the graph Laplacian has converged toward the Laplace Beltrami operator.
Again, this is left for a future contribution.

\subsection{}
\begin{mdframed}[style=comment]
A question that I would like to know personally, are you thinking about a future upgrade to non-radial kernels?
\end{mdframed}
We thought about it but are not sure that it would be beneficial.
%The main difficulty is that a graph is not intrinsically oriented.
%However, as the graph framework we are using only works with radial filters, it is not a simple question.
%A simple trick to make the network sensitive to direction is to add a per-channel fully connected layer.
%We are currently investigating other potential graph techniques that would lead to non-radial kernels.

\subsection{Details}
\begin{mdframed}[style=comment]
Finally, I noticed a few points in the text where there were typos or things which were difficult to follow and I have put a note of them below and tried to correct them as much as I can so that you don't have to put too much effort in - I can't wait to see this paper published. Sorry that the list is so extensive (I'm a bit of a stickler for details and trying to write papers concisely):
\end{mdframed}

Thank you very much for taking the time for making these corrections.
We found them very helpful in improving the quality of the paper.
We implemented most of them, see below.

@@

\begin{mdframed}[style=comment]
Convolutional Neural Networks (CNNs) have been proposed as an alternative analysis tool in cosmology thanks to their ability to automatically design relevant statistics to maximise the precision of the parameter estimation, while maintaining robustness to noise [9–16].

- There are many of us in the community who are not overly keen on "precision" of parameter estimation being the target for machine learning. Rather, we would prefer robustness of parameter estimation. It might be worth changing this sentence a bit so as not to upset some of the referenced authors ;)
\end{mdframed}

Precision, understood as the final size of the posterior distribution on the measured parameters (including systematic errors), is the key interest of cosmology analysis.
The robustness to noise is also crucial, as demonstrated in [9].
%We expanded the sentence to include this.
We added a footnote to specify this.
% \TK{not sure what do to here.. will people get upset? "Precision" is commonly used in cosmology}

\begin{mdframed}[style=comment]
When you have multiple = in equations could you split them on to separate lines? This would make the logical steps for following the equations a lot clearer.
\end{mdframed}

Thank you for the suggestion, normally we would adopt it, but the paper is already quite long, and breaking up the lines in equations would make it even longer. So we decided against it.

\begin{mdframed}[style=comment]
You change a lot between ``node'', ``vertex'' and ``pixel''. Could you stick to one of these? I think that ``vertex'' or ``pixel'' makes the most sense (``pixel'' due to the HEALPix pixelisation of the maps or vertex because of the use of the $v$ notation). Node is quite confusing in several places because there is an inherent connection between node and neuron in ML literature, and since there is swapping between the three different terms it's a little hard to keep track.
\end{mdframed}

We agree that using different terms for the same notion can confuse readers.
While ``node'' and ``vertex'' are synonyms in referring to the elements of a graph, a ``pixel'' is a different concept referring to a point of the (spherical) sampling.
A pixel can be represented by a vertex, but they are not the same.
As such, we decided to replace all occurrences of ``node'' with ``vertex'' but keep the two concepts of ``vertex'' and ``pixel'' separate.

% \nati{What do you think? This avoid repetition. But we could be more consistent...}
% \TK{Agreed. Do you want to make the change @nati?}
% \mdeff{Agreed. Node is the same as vertex. Pixel is not necessarily.}

\begin{mdframed}[style=comment]
You change between math font and text font for the names of your layers when inside and outside of equations. It would look a lot neater sticking with one choice. (Sorry, that seems quite petty and I don't mean it to be).
\end{mdframed}
We have made the paper consistent by using a math font for layer names.
%\nati{@all, when you read the paper, please check that I did not forget some of them. Every layer has been changed to math symbols as we use them like function. For example, SM becomes $SM$.}
%\mdeff{Agreed. I checked and fixed one missing.}

\begin{mdframed}[style=comment]
In Appendix C. you talk about Kroneker a lot, but earlier you talk about Kroneker $\delta$ (which is it's correct name). It's probably worth sticking with the full name.
\end{mdframed}
We have modified the manuscript according to the suggestion.

\begin{mdframed}[style=comment]
Sky maps are rotation invariant: rotating maps on the sphere doesn’t change their interpretation, as only the statistics of the maps are relevant.
- I think that you mean that rotating maps on the sphere doesn't change their interpretation *when* only the statistics of the maps are relevant. Say you wanted to compare the NGC against the SGC, then rotating the map would obviously change the results.
\end{mdframed}
True.
Equivariance is desired in such a case (rather than invariance).
We clarified the sentence.

@@

\begin{mdframed}[style=comment]
This kind of maps can be created using the gravitational lensing technique [see 29, for review]

- It should be "These" and not "This". "These kind of maps can be created using the gravitational lensing technique [see 29, for review]"
\end{mdframed}
Thank you for spotting this mistake. We updated the manuscript.

\begin{mdframed}[style=comment]
A CNN is composed of the following main building blocks [31]: (i) a convolution, (ii) a non-linearity, (iii) a down-sampling operation, (iv) a pooling operation, and (v), optionally, normalization.

- I think that this is generalising a CNN a bit too much. In my mind (and in my personal uses of CNNs) steps (iii) and (iv) are not always necessary either. Perhaps you want to say that you wish to tackle common tasks which often arise in CNNs: i) convolution, (ii) non-linearity, (iii) down-sampling operation, (iv) pooling operation, and (v), optionally, normalisation.
\end{mdframed}
We agree that (iii) and (iv) are optional and updated the sentence accordingly.

@@

- Might it be more concise to write "The resolution changes as $N_{pix} = N^2_{side}\times12$ such that $N_{pix} = 48$ for $N_{side} = 2$ and $N_{pix} = 192$ with $N_{side}=3$."
\end{mdframed}
Thank you for the suggestion, we updated the text.

\begin{mdframed}[style=comment]
... is a measure of the variation of the eigenvector $\bf{u}_i$ is on the graph defined by the Laplacian L.

 - I don't think the "is" after $\bf{u}_i$ should be there.
\end{mdframed}
This has been corrected.

\begin{mdframed}[style=comment]
Given the convolution kernel $h:\mathbb{R}_+ \to \mathbb{R}$, a signal $f\in\mathbb{R}^{N_{pix}}$ on the graph is filtered as

- You don't mention what $\mathbb{R}_+$ is or why $h$ defines that map.
\end{mdframed}
We are not sure what the reviewer means here.
$\mathbb{R}_+$ is the ensemble of the real positive numbers and $h$ is a function that maps any positive real number to a real number.
We think that $\mathbb{R}_+$ is a common enough notation to not require being explicitly defined.

%\nati{This seems clear to me. Specifying what $\mathbb{R}_+$ is would be a bit too much I believe. @all, is it also for you?}
%\mdeff{That is also quite clear to me.}

### 2019-03-04 scnn.tex
comment 2.14 (details) part 1

So far these algorithms have mostly been demonstrated on Euclidean domains, such as images.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this domain, while taking care of the necessary irregular sampling.
Moreover, the designed convolution and resulting neural network should possess the following three key characteristics.
First, the convolution should be equivariant to rotation, meaning that a rotation of the input implies the same rotation of the output.
\nati{We consider the sky maps rotationial invariant as in our case rotating maps on the sphere doesn't change their interpretation since only the statistics of the maps are relevant.}
Depending on the task, we want the CNN to be either equivariant or invariant to rotation.
Second, to be able to train the network in reasonable time, the convolution has to be computationally efficient.
Third, a CNN should work well on parts of the sphere, as many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area (see \figref{example_maps} for example maps).

@@

A CNN is composed of the following main building blocks~\citep{lecun1998cnn}:
(i) a convolution,
(ii) a non-linearity.
Optionally, some CNN also have
(iii) a down-sampling operation,
(iv) a pooling operation, and
(v) a normalization.\footnote{Batch normalization has been shown to help training \citep{ioffe2015batchnorm}. We verified this experimentally in our setting as well.}
Our architecture is depicted in~\figref{architecture} and discussed in greater details in \secref{architecture}. As operations (ii) and (v) are point-wise, they do not depend on the data domain. The pooling operation is simply a permutation invariant aggregation function which does not need to be adapted either. The convolution and down-sampling operations, however, need to be generalized from Euclidean domains to the sphere.

On regular Euclidean domains, such as 1-dimensional time series or 2-dimensional images, a convolution can be efficiently implemented by sliding a localized convolution kernel (for example a patch of $5 \times 5$ pixels) in the signal domain.
\nati{Likewise, down-sampling can be achieved by keeping one pixel every $n$ pixels with $n=2$ being very common.}
Because of the irregular sampling, there is no straightforward way to define a convolution on the sphere directly in the pixel domain; convolutions are most often performed using spherical harmonics.
In our method we also use the spectral domain to define the convolution.
The implementation, however, does not need a direct access to the spectrum, which is computationally more efficient (see \secref{efficient_convolution}).
The gist of our method is to define the convolution operation on a sphere using a graph, and the down-sampling operation using a hierarchical pixelisation of the sphere.

@@

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal-sized pixels (right-most sphere in \figref{pooling}).
% As each pixel is subdivided in four, the second coarser resolution is $N_{pix} = N_{side}^2 \times 12 = 2^2 \times 12 = 48$ pixels (middle sphere in \figref{pooling}), the third is $N_{pix} = N_{side}^2 \times 12 = 4^2 \times 12 = 192$ pixels, etc., where $N_{side} = 1, 2, 4, 8, \ldots$ is the grid resolution parameter.
The resolution changes as $N_{pix} = N^2_{side}\times12$ such that $N_{pix} = 48$ for $N_{side} = 2$ and $N_{pix} = 192$ with $N_{side}=3$.
High-resolutions HEALPix maps easily reach millions of pixels.

@@

Fully convolutional networks (FCNs) have been introduced by  \citep{long2015fcn} and are mostly used for the semantic segmentation of images, a dense classification task.
An example FCN for dense regression is
\begin{equation*}
	\Y = NN_\theta(\X) = (GC \circ \sigma \circ BN \circ GC)(\X) \in \R^{N_{pix} \times F_{out}},
\end{equation*}
where $\circ$ denotes composition, i.e., $(f \circ g)(\cdot) = f(g(\cdot))$.
If P layers are used, they have to be inverted via up-sampling in an encoder-decoder architecture.

@@

In \figref{gaussian_filters_visualization}, we further leverage this example to present the three visualizations of a HEALPix graph convolution kernel.
We remember that the convolution kernel resulting from the heat diffusion is defined as $K_t(x)=e^{-\tau t x}$. For our (top) visualization, we simply plot this function, i.e., we evaluate $K_t$
at the graph eigenvalues $\text{diag}(\bLambda)$. Note that, the lower eigenvalues $\lambda$ correspond to the lower frequencies, i.e., the spectral mode with less variation.
As a second visualization (middle), we observe the kernel directly on the sphere. As the kernel $K_t$ is defined in the graph spectral domain, we convolve a Kroneker $\delta$  to obtain a representation on the sphere. We then plot the result with a gnomonic projection.
\nati{For our third visualization (bottom), our goal is to plot the section of a filter. This visualization might be convenient as we theoretically have spherical filters (up to the irregularities of the sampling). The idea is to plot the section of the convolved Kroneker $\delta$.}
To do so, we convolve a Kroneker $\delta$ on the vertices along the equator and use the equator for our section.
Note that, because of the small irregularities in the HEALPix sampling, the second and the third methods are subject to small variations depending on the chosen position of the Kroneker $\delta$ .

>>>>

So far these algorithms have mostly been demonstrated on Euclidean domains, such as images.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this domain, while taking care of the necessary irregular sampling.
Moreover, the designed convolution and resulting neural network should possess the following three key characteristics.
First, the convolution should be equivariant to rotation, meaning that a rotation of the input implies the same rotation of the output.
Sky maps are rotation equivariant: rotating a map on the sphere doesn’t change its interpretation.
Depending on the task, we want the CNN to be either equivariant or invariant to rotation.\footnote{When only the statistics of the maps are relevant, they are rotation invariant.}
Second, to be able to train the network in reasonable time, the convolution has to be computationally efficient.
Third, a CNN should work well on parts of the sphere, as many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area (see \figref{example_maps} for example maps).

@@

A CNN is composed of the following main building blocks~\citep{lecun1998cnn}:
(i) a convolution,
(ii) a non-linearity, and, optionally,
(iii) a down-sampling operation,
(iv) a pooling operation, and
(v) a normalization.\footnote{Batch normalization has been shown to help training \citep{ioffe2015batchnorm}. We verified this experimentally in our setting as well.}
Our architecture is depicted in~\figref{architecture} and discussed in greater details in \secref{architecture}. As operations (ii) and (v) are point-wise, they do not depend on the data domain. The pooling operation is simply a permutation invariant aggregation function which does not need to be adapted either. The convolution and down-sampling operations, however, need to be generalized from Euclidean domains to the sphere.

On regular Euclidean domains, such as 1-dimensional time series or 2-dimensional images, a convolution can be efficiently implemented by sliding a localized convolution kernel (for example a patch of $5 \times 5$ pixels) in the signal domain.
Because of the irregular sampling, there is no straightforward way to define a convolution on the sphere directly in the pixel domain; convolutions are most often performed using spherical harmonics.
In our method we also use the spectral domain to define the convolution.
The implementation, however, does not need a direct access to the spectrum, which is computationally more efficient (see \secref{efficient_convolution}).

Down-sampling is achieved on regular Euclidean domains by keeping one pixel every $n$ pixels in every dimension.
That is again not a suitable strategy on the sphere because of the irregular sampling.

The gist of our method is to define the convolution operation on a sphere using a graph, and the down-sampling operation using a hierarchical pixelisation of the sphere.

@@

The lowest possible resolution is given by the base partitioning of the surface into $N_{pix} = 12$ equal-sized pixels (right-most sphere in \figref{pooling}).
The resolution changes as $N_{pix} = 12 N^2_{side}$ such that $N_{pix} = 48$ for $N_{side} = 2$ and $N_{pix} = 192$ for $N_{side} = 3$.
High-resolutions maps easily reach millions of pixels.

@@

Fully convolutional networks (FCNs) have been introduced by  \citep{long2015fcn} and are mostly used for the semantic segmentation of images, a dense classification task.
An example FCN for dense regression is
\begin{equation*}
	\Y = NN_\theta(\X) = (GC \circ \sigma \circ BN \circ GC)(\X) \in \R^{N_{pix} \times F_{out}},
\end{equation*}
where $\circ$ denotes composition, i.e., $(f \circ g)(\cdot) = f(g(\cdot))$.
If $P$ layers are used, they have to be inverted via up-sampling in an encoder-decoder architecture.

@@

In \figref{gaussian_filters_visualization}, we further leverage this example to present the three visualizations of a HEALPix graph convolution kernel.
We remember that the convolution kernel resulting from the heat diffusion is defined as $K_t(x)=e^{-\tau t x}$. For our (top) visualization, we simply plot this function, i.e., we evaluate $K_t$
at the graph eigenvalues $\text{diag}(\bLambda)$. Note that, the lower eigenvalues $\lambda$ correspond to the lower frequencies, i.e., the spectral mode with less variation.
As a second visualization (middle), we observe the kernel directly on the sphere. As the kernel $K_t$ is defined in the graph spectral domain, we convolve a Kronecker $\delta$ to obtain a representation on the sphere. We then plot the result with a gnomonic projection.
\nati{For our third visualization (bottom), our goal is to plot the section of a filter. This visualization might be convenient as we theoretically have spherical filters (up to the irregularities of the sampling). The idea is to plot the section of the convolved Kronecker $\delta$.}
To do so, we convolve a Kronecker $\delta$ on the vertices along the equator and use the equator for our section.
Note that, because of the small irregularities in the HEALPix sampling, the second and the third methods are subject to small variations depending on the chosen position of the Kronecker $\delta$.
